---
title: "TRBL Acoustic Methods"
output:
  word_document: default
  html_document: default
date: "2023-02-26"
---

## Using records from 2019 data to evaluate how call type, month, and site affect odds of correct detection

## Using results from manually analyzed data and pattern matching algorithms from semi-automated data

Note that site Creed Road = Denverton Creek

```{r}
# load in the data
df <- read.csv("~/Desktop/TRBL Bioacoustics/MethodsData.csv", stringsAsFactors=F)  

# convert call type results to "long" format
ldf <- reshape(df, direction="long", varying = list(7:11), v.names = "Evaluation", timevar = "CallType", 
             times = names(df[,c(7:11)]))
ldf <- ldf[,which(names(ldf)!="id")]

# we have some detections labeled as missed which are technically false negatives (FN)
ldf$Evaluation <- ifelse(ldf$Evaluation=="missed","FN",ldf$Evaluation)

# now we need to add a response variable
# 1 = correctly detected & 0 = incorrectly detected
ldf$Response <- ifelse(ldf$Evaluation %in% c("TP","TN"),1,0)
```

## now that the data is properly formatted we can model the data

```{r}

# test if there are differences in the odds of a TP or TN among call types
mdl<-glm("Response ~ CallType + hour + site", data=ldf, family="binomial")

# see the results
summary(mdl)

```

There is no evidence of an effect of hour on accuracy of detection so we remove hour

```{r}

# testing difference in the odds of detection using call type, month, site
mdl<-glm("Response ~ CallType + month + site",data=ldf,family="binomial")
summary(mdl)

```

There is a decline in accuracy of detection with month

```{r}

# visualize decline of detection with month
library(ggplot2)
pdfa<-aggregate(Response~month+site,ldf,sum)
pdfb<-aggregate(Response~month+site,ldf,NROW)
names(pdfb)<-c("month","site","numTests")
pdf<-merge(pdfa,pdfb,by=c("month","site"))
pdf$accRate<-pdf$Response/pdf$numTests
ggplot(pdf,aes(x=month,y=accRate)) + geom_bar(stat="identity",aes(fill=site),width=0.8,position="dodge") +
  labs(x="Month",y="Accuracy rate",fill="Location") + theme_bw()

```


The decline of accuracy of detection with month is mainly from Rush Ranch 
Adding an interaction with site

```{r}

# Need to make month a factor, not numeric
# Effect of month was not linear so let it vary by making it a factor
ldf$Month<-ifelse(ldf$month==2,"Feb",ifelse(ldf$month==3,"Mar",ifelse(ldf$month==4,"Apr",ifelse(ldf$month==5,"May",ifelse(ldf$month==6,"Jun","Jul")))))

mdl<-glm("Response ~ CallType + Month*site",data=ldf,family="binomial")
summary(mdl)

# Output of table to include in manuscript
library(stargazer)
stargazer(mdl, type = "html", single.row = TRUE,
          out="mdl.html")

```

Reference call is female song Reference site is Creed Road/Denverton Creek Reference month is April

```{r}

# exponentiate the intercept (female song) you get these odds:
exp(-0.48489)

# the odds of a correct detection for female (i.e., TP or TN) are this amount lower than getting it wrong 
1-exp(-0.48489)

```

```{r}

# the odds of a correct detection of male song
ms <-  -0.48489 + 0.12791
exp(ms)

# the odds of a correct detection for nestling calls
nc <- -0.48489 + 1.49592
exp(nc)

# the odds of a correct detection for fledgling calls
fc <- -0.48489 + 1.29804
exp(fc)

```
Look at how odds of correct detection varies by site

```{r}

# Hay Landfill odds
exp(-1.04398)
  # Hay Landfill odds of correct detection are lower than at Denverton Creek

# Rush Ranch odds
exp(2.45936)
  # Rush Ranch enhances odds of correct detection 

```

```{r}

# what is most important in affecting the accuracy rate
anova(mdl)
  # site accounts for most deviance

```

